---
title: "p8105_hw2_lcc2181"
output: github_document
---

## Problem 1

Loading the packages needed for Problem 1.

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

Reading and cleaning the Mr. Trash Wheel sheet. This includes improving variable names, removing rows that do not have dumpster-specific data, and rounding the number of sports balls to the nearest integer.

```{r}
trash_wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
    sheet = "Mr. Trash Wheel", 
    range = "A2:N408") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Reading and cleaning precipitation data for 2018. This includes omitting rows that do not have precipitation data for a specific month and adding a variable to indicate the year associated with the data.

```{r}
precip_2018_df = 
  read_excel ("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total, month) %>% 
  mutate(year = 2018)
```

Reading and cleaning precipitation data for 2019. This includes omitting rows that do not have precipitation data for a specific month and adding a variable to indicate the year associated with the data.

```{r}
precip_2019_df = 
  read_excel ("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total, month) %>% 
  mutate(year = 2019)
```

Combining precipitation data sets using a full join to include all precipitation data for 2018 and 2019. Months are converted from a numeric to character variable listing the full name of each month.

```{r}
precipitation_df = 
  full_join(precip_2018_df, precip_2019_df) %>%
  mutate(month = month.name[as.numeric(month)])
```

The number of observations in the cleaned Mr. Trash Wheel data is `r nrow(trash_wheel_df)`. There are `r ncol(trash_wheel_df)` variables in the cleaned Mr. Trash Wheel data. Examples of key variables in the cleaned Mr. Trash Wheel data are dumpster, month, year, date, and homes_powered. The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trash_wheel_df, year == 2019), sports_balls))`. The number of observations in the combined precipitation data set is `r nrow(precipitation_df)`. There are `r ncol(precipitation_df)` variables in the combined precipitation data set, which are `r names(precipitation_df)`. The total precipitation in the year 2018 was `r sum(pull(filter(precipitation_df, year == 2018), total))`.

## Problem 2

Cleaning data in pols-months.csv. Separating variable mon into three integer variables -- year, month, and day. Months are converted into character variable containing month name. Since presidents are either gop or dem, the president variable is created based on the data from the prez_dem column, assuming that if the president is not democratic, the president is republican. Finally, the prez_gop and prez_dem varaibles, as well as the day variable, are removed.

```{r}
pols_month_df =
  read_csv("data/pols-month.csv", col_types = "cdddddddd") %>% 
  separate(mon, into= c("year","month","day"), sep= "-", convert = TRUE) %>% 
  mutate(
    month = month.name[as.numeric(month)],
    president = recode(prez_dem, `1`= "dem", `0`= "gop" )
  ) %>% 
  select(-prez_gop, -prez_dem, -day)
```

Cleaning data in snp.csv. Dates are converted to a format with a four digit year, and then separated into three integer variables -- year, month, and day. Months are converted into character variable containing month name. Year and month are organized to be the two leading columns, and day is removed.

```{r}
snp_df =
  read_csv("data/snp.csv", col_types = "cd") %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year","month","day"), sep = "-", convert = TRUE) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  relocate(year,month) %>% 
  select(-day)
```

Tidying data in unemployment.csv. Using pivot_longer to switch from wide to long format. Months are re-coded  from abbreviated to full to ensure consistency with the other data sets. Names are cleaned to ensure variable names are lowercase to be consistent with other the data sets.

```{r}
unemployment_df =
  read_csv("data/unemployment.csv", col_types = "idddddddddddd") %>% 
  pivot_longer(Jan:Dec,
      names_to = "month",
      values_to = "unemployment"
  ) %>% 
  mutate(month = recode(month, 
        "Jan" = "January", "Feb" = "February", "Mar" = "March", "Apr"=    
        "April", "May" = "May", "Jun" = "June", "Jul" = "July", "Aug" =  
        "August", "Sep" =  "September", "Oct" = "October", "Nov" = 
        "November", "Dec" = "December")) %>% 
  janitor::clean_names()
```

Joining data sets by merging snp_df into pols_month_df, and then merging unemployment into result.

```{r}
pols_snp_df = 
  left_join(pols_month_df, snp_df, by = c("year", "month"))

fivethirtyeight_df =
  left_join(pols_snp_df, unemployment_df, by = c("year","month"))
```

The pols-month data contained information about the number of republican and democratic governors, senators, and representatives on the associated date as well as whether the president was republican and democratic. The snp data contained information about the closing values of the S&P stock index on the date associated. The unemployment data contained measurements of the percentage of unemployment by month and year. The resulting merged data set has `r nrow(fivethirtyeight_df)` rows and `r ncol(fivethirtyeight_df)` columns. It ranges from the years `r min(pull(fivethirtyeight_df, year))` to `r max(pull(fivethirtyeight_df, year))`. The key variables in the resulting data set include year, month, president, close, and unemployment.

## Problem 3

Loading and tidying data from Popular_Baby_Names.csv

```{r}
baby_names_df =
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    ethnicity = recode(ethnicity, 
    "BLACK NON HISP" = "BLACK NON HISPANIC", 
    "WHITE NON HISP" = "WHITE NON HISPANIC", 
    "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER")
  ) %>% 
  mutate(
    childs_first_name = str_to_title(childs_first_name)
  ) %>% 
  distinct()
```

Table showing rank in popularity of name "Olivia" as female baby name over time

```{r}
filter(baby_names_df, gender == "FEMALE", childs_first_name == "Olivia") %>%
  pivot_wider(
   id_cols = c("ethnicity"),
   names_from = "year_of_birth",
   values_from = "rank"
   ) %>% 
  knitr::kable()
```

Table showing most popular name among male children over time

```{r}
filter(baby_names_df, gender == "MALE", rank == 1) %>% 
  pivot_wider(
    id_cols = c("ethnicity"),
    names_from = "year_of_birth",
    values_from = "childs_first_name"
    ) %>% 
  knitr::kable()
```

Scatter plot showing number of children with name against rank in popularity of that name among male, white non-hispanic children born in 2016

```{r}
filter(baby_names_df, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016)%>% 
ggplot(aes(x = rank, y = count)) + geom_point()
```

