---
title: "p8105_hw2_lcc2181"
output: github_document
---

## Problem 1

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
```

Reading and cleaning the Mr. Trash Wheel sheet.

```{r}
trash_wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet = "Mr. Trash Wheel", 
  range = "A2:N408") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls, digits = 0))
```

Reading and cleaning precipitation data for 2018.

```{r}
precip_2018_df = 
  read_excel ("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  drop_na(month) %>% 
  mutate(year = 2018)
                 
```

Reading and cleaning precipitation data for 2019.

```{r}
precip_2019_df = 
  read_excel ("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  drop_na(month) %>% 
  mutate(year = 2019)
```

Combining precipitation data sets.

```{r}
precipitation_df = 
  full_join(precip_2018_df, precip_2019_df) %>%
  mutate(month = recode(month, 
        `1`= "January", `2`= "February", `3`= "March", `4`= "April",
        `5`= "May", `6`= "June", `7`= "July", `8`= "August", 
        `9`=  "September", `10`= "October", `11`= "November", 
        `12`= "December")
    )
```

The number of observations in the cleaned Mr. Trash Wheel data is `r nrow(trash_wheel_df)`. There are `r ncol(trash_wheel_df)` variables in the cleaned Mr. Trash Wheel data. Examples of key variables are dumpster, month, year, date, and homes_powered. The median number of sports balls in a dumpster in 2019 was `r median(pull(filter(trash_wheel_df, year == 2019), sports_balls))`. The number of observations in the cleaned 2018 and 2019 precipitation data is `r nrow(precipitation_df)`. There are `r ncol(precipitation_df)` variables in the precipitation data, which are `r names(precipitation_df)`. The total precipitation in 2018 was `r sum(pull(filter(precipitation_df, year == 2018), total))`.

